---
title: "Adv. Data Cleaning"
author: "Amanda Mae Woodward"
date: "4/7/2022"
output: html_document
---

```{r loadData}
library(tidyverse)
data(billboard)
View(billboard)
```

# Things to do to clean data (make it make sense)

- missing values- a lot... 
- remove NA, long format

```{r checkMissing}
summary(is.na(billboard)==TRUE)  #early columns no missing; late columns missing everything
```

```{r removingMissing}
#na.omit- remove rows with missing data 
billboard2<- na.omit(billboard) #bad idea, gets rid of all of our data
```

Use long format
```{r tolongFormat}
# cols= are the things we want to be in long format; everything else is ID variable
billboardLong<- billboard %>% pivot_longer(cols=wk1:wk76, names_to= "week", values_to= "position") #this is hard coded names; means same columns will be used regardless of cleaning

billboardL3<- billboard %>% pivot_longer(cols= c(colnames(billboard)[4:79]), names_to= "week", values_to= "position", values_drop_na = TRUE) #soft code
```

```{r}
billboardLong2<- na.omit(billboardLong)
```

- add a column for length of time on top 100 (song and artist)
steps to do this: 
- add a column 
- calculate length of time 
- in wide data: count the number of columns (in long data this will be #of rows)
How can we do this? 
- count function 
```{r}
artistList<-billboardLong %>% group_by(artist, track) %>% tally()
artistList
#why are things popping up 76 times? 
summary(as.factor(artistList$n)) #that's not right. figure out later
#fix 
RealartistList<-billboardL3 %>% group_by(artist, track) %>% tally()
RealartistList
```

```{r addtoData}
left_join(billboard, RealartistList)#this will combine data sets of the same length

longDat4<-(left_join(billboardL3, RealartistList)) # this works to join datasets that are different sizes. When using with own data, double check that this works 

#change column name of n
colnames(longDat4)[6]<- "chartFreq"

#names(longDat4)[names(longDat4) == 'n'] <- 'chartfreq' or this
```

- multiple songs; may want to know about artist (artist level info)
- use same logic as above

- splitting by year  
date entered column has year, month, day
```{r}
#separate into year, month, day columns
newDate<-billboard %>%  separate(col= date.entered, into= c("year", "month", "day"), sep= "-")
```

```{r}
sortedDat<- newDate %>% arrange(year)
```

- sort by/arrange by frequency of time on list artist 
- whether or not song entered billboard top 100 on birthday 
